Designing and Developing a Personalised Recommender System

### 2021-02-11 paper/paper.tex
renamed files and cleaned up comments

\subsection{Data description}
The data for my recommender system was generated from the Yelp Open Dataset \cite{YelpDataset}. 
This is a subset of all the businesses, reviews, and user data on Yelp, formatted as JSON files. 
The dataset includes approximately 8,000,000 reviews of 200,000 businesses by 2,000,000 users, as well as 
additional Covid-19 related features of each business such as whether they now offer takeout or delivery. 
The dataset also includes approximately 200,000 photos but I these are not relevant to my application, since 
the interface is command line only, so they were discarded. 
Every businesses in the dataset is in 1 of 10 metropolitan areas in the USA and Canada 
(Montreal, Calgary, Toronto, Pittsburgh, Charlotte, Urbana-Champaign, Phoenix, Las Vegas, Madison, and Cleveland) \cite{YelpDataset}.

\subsection{Data preparation and feature selection}
The business data was filtered down to businesses that contain 'Restaurants' as one of their listed 
categories
The ratings data was 
(20,565 restaurants)
Prepare data
Of the 209,393 businesses in the dataset, 168,903 are still open. 
There are 43,965 restaurants which are open and contain 'Restaurants' as one of the categories. 
I then picked the top 12 most popular categories of restaurant: American, Mexican, Italian, Chinese, 
Seafood, Japanese, Canadian, Mediterranean, Indian, Thai, Middle Eastern, Vietnamese (in that order). 
When filtered for restaurants which fall into at least one of these categories, 
20,565 had at least one check-in in 2019, but 333 of these were temporarily closed due to covid but should be 
open now, and 1 is still closed until June 2021. 
19,862 restaurants had a review written about them in 2019, there were 503,979 reviews written in 2019 about these restaurants. 
There are 2,956,316 reviews of these restaurants. 
481,073 tips for these restaurants. 
1,968,703 users. 
None of these restaurants are temporarily closed according to the covid features data. 
20,573 restaurants listed with covid features, 4 restaurants which make up 8 duplicate entries 
(4 have 3 entries), the only difference between their entries being in the 
"Virtual Services Offered" section. 
Allow to pick user ID of 10 users with most reviews made, to avoid cold start problem
4 users have over 10,000 reviews, the most is 14,455 (Victor, 8k3aO-mPeyhbR5HUucA5aA). 
I actually want to select the most reviews of the places I am accepting, restaurants. 
When taking the most reviews of restaurants in teh dataset, a threshold of 600 reviews gives a list of 10 users, 
with the most being 1747. 
I then checked each of these for how many different restaurants they had reviewed and the greatest number of times 
they had reviewed a single restaurant to check that none of them were just users who reviewed a small number of 
restaurants a huge amount of times each. 
All 10 users passed this criteria, with even the least active user having reviewed 496 unique restaurants, which will 
certainly avoid any cold-start problems. 
I split the businesses into 10 groups, by the city which they are closest to, I did this by using latitude and 
longitude rather than State or City since this does not give actual distance and restaurants close to a border 
could then be mis-categorised. 
Since there are 4,223,201 reviews of restaurants I filtered them by removing any which had zero 'useful' votes. 
When filtered to reviews with at least 1 'useful' vote, there are 1,735,732 and when at least 2 votes, 836,929
At least 3 votes, 462,412, at least 5 votes 190,902. 
51,737 with at least 10 votes

@@

\subsection{Recommendation techniques/algorithms}
The first recommender system is collaborative filtering.
I am using item-based collaborative filtering, since the number of users is larger than the number of items 
(1,968,703 users vs 209,393 businesses in the entire dataset), this gives greater accuracy of predictions. 
In this case the list of available items does not change, so a user-based method would be unnecessary. 
One of the requirements for this system is justifiability, and item-based methods make it much easier to 
explain why certain predictions were made. 
Item-item collaborative filtering looks for restaurants that are similar, in terms of how people rate them, 
to the restaurants that the user has already rated and recommend the most similar restaurants. 
Filtering reviews by city and then using CF won't work because the active user may not have ever reviewed a restaurant 
in that city, so there is no way to assess similar restaurants to ones they have already rated. 
Instead I carried out CF on all high quality reviews and produce an ordering of recommendations, which I then filter 
by city. 
Mention Netflix Prize and proposition of SVD, use lecture slides explanation and surprise documentation.

@@

\subsection{Recommendation algorithm}
I used a model-based approach to Collaborative Filtering based on matrix factorisation. 
The specific model is Single Value Decomposition (SVD), which was popularised by Simon Funk during the Netflix Prize. 
In my program, I fit the algorithm on a training set, and then for every restaurant of the specified category 
and in the specified city, I predict what the user would rate that restaurant. 
Then I sort those restaurants by their score and take the top 8 to present to the user.

@@

\subsection{Further developments}
What could I do in the future \cite{5284958}.

>>>>

\subsection{Data description}
The data for my recommender system was generated from the Yelp Open Dataset \cite{YelpDataset}. 
This is a subset of all the businesses, reviews, and user data on Yelp, formatted as JSON files. 
The dataset includes approximately 8,000,000 reviews of 200,000 businesses by 2,000,000 users, as well as 
additional Covid-19 related features of each business such as whether they now offer takeout or delivery. 
There are 3 sections of the dataset, with 5 JSON files in the main dataset, 1 in the photos section, and 
1 in the covid section. 
The $business.json$ file contains business data including location data, attributes, and categories. 
$review.json$ contains full review text data including the $user\_id$ that wrote the review and the $business\_id$ 
that the review is written for. 
The file $user.json$ holds every user's first name along with other data adn metadata which I don't use \cite{YelpDataset}. 
The data files $checkin.json$ and $tip.json$ are not used in my system. 
The dataset also includes approximately 200,000 photos but these are not relevant to my application, since 
the interface is command line only, so they were discarded. 
The covid related data from $covid\_features.json$ contains additional information about each business with regards to 
the measures they have taken due to the pandemic, such as offering delivery or takeout options or a 
special message for customers. 
Every business in the dataset is in 1 of 10 metropolitan areas in the USA and Canada 
(Montreal, Calgary, Toronto, Pittsburgh, Charlotte, Urbana-Champaign, Phoenix, Las Vegas, Madison, and Cleveland) \cite{YelpDataset}.

\subsection{Data preparation and feature selection}
The dataset is provided in JSON format, but in such a way that every entry is its own JSON on its own line, 
which does not lend well to fast lookups by the main program. 
Therefore it was necessary to convert every file I used in the system into a typical JSON file, such that it could 
be loaded as a python dictionary with the lookup key being either the uniquely identifying $business\_id$ or 
$user\_id$ depending on the case. 
This allows for extremely fast data lookups in the system. 
It was of course also necessary to filter the list of all Yelp businesses down to just the restaurants. 
I further narrowed the list of restaurants by selecting only those which fit into one or more of the top 12 
categories in terms of number of restaurants (American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, 
Mediterranean, Indian, Thai, Middle Eastern, Vietnamese, in that order). 
Since the user selects one of these 12 options for categories of restaurants they want to find, they will 
never be recommended a restaurant which doesn't fit at least one of these categories. 
I also ensured that every restaurant was still open and had at least 1 check-in in 2019 (the last year that 
data was available). 
This left 20,565 restaurants which could be recommended to the user, which needed to be split into their respective 
metropolitan areas by using the latitude and longitude values of the businesses and of the centres of each city. 
The list of reviews pertaining to these restaurants was then filtered down
I also split by nearest city, and filtered only useful reviews
This was done using the $categories$ attribute of the $business$ JSON file.

The business data was filtered down to businesses that contain 'Restaurants' as one of their listed 
categories
getIDs
filterReviews
lastCheckin
recentCheckin
toCSV
CovidDuplicates
getUserIDs
mostActiveUsers
numberRestaurantReviews
mostReviews
filterMostActiveReviews
uniqueRestaurants
dropText
splitCities
splitIDs
filterUsefulReviews
filterRecentReviews
splitReviews
getCategories
formatCovidFeatures
closedCovid

The ratings data was 
(20,565 restaurants)
Prepare data
Of the 209,393 businesses in the dataset, 168,903 are still open. 
There are 43,965 restaurants which are open and contain 'Restaurants' as one of the categories. 
I then picked the top 12 most popular categories of restaurant: American, Mexican, Italian, Chinese, 
Seafood, Japanese, Canadian, Mediterranean, Indian, Thai, Middle Eastern, Vietnamese (in that order). 
When filtered for restaurants which fall into at least one of these categories, 
20,565 had at least one check-in in 2019, but 333 of these were temporarily closed due to covid but should be 
open now, and 1 is still closed until June 2021. 
19,862 restaurants had a review written about them in 2019, there were 503,979 reviews written in 2019 about these restaurants. 
There are 2,956,316 reviews of these restaurants. 
481,073 tips for these restaurants. 
1,968,703 users. 
None of these restaurants are temporarily closed according to the covid features data. 
20,573 restaurants listed with covid features, 4 restaurants which make up 8 duplicate entries 
(4 have 3 entries), the only difference between their entries being in the 
"Virtual Services Offered" section. 
Allow to pick user ID of 10 users with most reviews made, to avoid cold start problem
4 users have over 10,000 reviews, the most is 14,455 (Victor, 8k3aO-mPeyhbR5HUucA5aA). 
I actually want to select the most reviews of the places I am accepting, restaurants. 
When taking the most reviews of restaurants in teh dataset, a threshold of 600 reviews gives a list of 10 users, 
with the most being 1747. 
I then checked each of these for how many different restaurants they had reviewed and the greatest number of times 
they had reviewed a single restaurant to check that none of them were just users who reviewed a small number of 
restaurants a huge amount of times each. 
All 10 users passed this criteria, with even the least active user having reviewed 496 unique restaurants, which will 
certainly avoid any cold-start problems. 
I split the businesses into 10 groups, by the city which they are closest to, I did this by using latitude and 
longitude rather than State or City since this does not give actual distance and restaurants close to a border 
could then be mis-categorised. 
Since there are 4,223,201 reviews of restaurants I filtered them by removing any which had zero 'useful' votes. 
When filtered to reviews with at least 1 'useful' vote, there are 1,735,732 and when at least 2 votes, 836,929
At least 3 votes, 462,412, at least 5 votes 190,902. 
51,737 with at least 10 votes

@@

\subsection{Recommendation techniques/algorithms}
The first recommender system is collaborative filtering. reference https://surpriselib.com/
I am using item-based collaborative filtering, since the number of users is larger than the number of items 
(1,968,703 users vs 209,393 businesses in the entire dataset), this gives greater accuracy of predictions. 
In this case the list of available items does not change, so a user-based method would be unnecessary. 
One of the requirements for this system is justifiability, and item-based methods make it much easier to 
explain why certain predictions were made. 
Item-item collaborative filtering looks for restaurants that are similar, in terms of how people rate them, 
to the restaurants that the user has already rated and recommend the most similar restaurants. 
Filtering reviews by city and then using CF won't work because the active user may not have ever reviewed a restaurant 
in that city, so there is no way to assess similar restaurants to ones they have already rated. 
Instead I carried out CF on all high quality reviews and produce an ordering of recommendations, which I then filter 
by city. 
Mention Netflix Prize and proposition of SVD, use lecture slides explanation and surprise documentation.

@@

\subsection{Recommendation algorithm}
The idea is to find restaurants in a new city which are similar to those in another city that the user likes, just like Entree. 
I used a model-based approach to Collaborative Filtering based on matrix factorisation. 
The specific model is Single Value Decomposition (SVD), which was popularised by Simon Funk during the Netflix Prize. 
In my program, I fit the algorithm on a training set, and then for every restaurant of the specified category 
and in the specified city, I predict what the user would rate that restaurant. 
Then I sort those restaurants by their score and take the top 8 to present to the user.

@@

\subsection{Further developments}
What could I do in the future \cite{5284958}. 
Implement the idea of particular categories being closer some than others. 
Specify price range and/or consider price when comparing similarities of restaurants.
